---
title: 'Reproducible Research: Peer Assessment 1'
output: md_document
---

###Loading and preprocessing the data


```{r}
library(lubridate)
activity_data <- read.csv("./activity.csv")
activity_data$data <- as.Date(activity_data$date, format="%Y-%m-%d")
activity_data$weekday <- wday(activity_data$date, label=T,abbr=T)
activity_data$weekday <- as.factor(activity_data$weekday)
activity_data$interval <- as.factor(activity_data$interval)
days <- as.factor(activity_data$date)
```


###What is mean total number of steps taken per day?


**Calculate the total number of steps taken per day and create a histogram**
```{r}
total_steps <- aggregate(steps ~ date, activity_data, sum)
hist(total_steps$steps, xlab= "Total Steps Taken per Day", col="blue", breaks = 10)
```

**Calculate the mean and median of the total number of steps taken per day**
```{r}
mean(total_steps$steps)
median(total_steps$steps)
```


###What is the average daily activity pattern?

```{r}
interval_split <- split(x=activity_data,f=activity_data$date)
interval_split <- interval_split[[1]]['interval']

library(lattice)
avg_steps_interval <- tapply(activity_data$steps, activity_data$interval, mean, na.rm=T )

xyplot(avg_steps_interval ~ interval_split,
       type="l",
       ylab="Mean Steps",
       xlab="Intervals",
       las=2,
       par.settings=list(layout.heights=list(top.padding=3, bottom.padding=5)))
       
names(which.max(x=avg_steps_interval))
```

**Total number of missing values**
```{r}
average <- tapply(activity_data$steps, activity_data$weekday, mean, na.rm=T )
total_NA <- activity_data[is.na(activity_data$steps),]
length(total_NA$steps)

days <- as.factor(activity_data$date)
X <- split(activity_data, days)

for(i in 1:length(X))
{
  for(j in 1:length(X[[i]][,'steps']))
  {
    if(is.na(X[[i]][j,'steps']))
    {
      new_day <- X[[i]][j,'date']
      X[[i]][j,'steps'] <- average[wday(new_day)]
    }  
  }
}

new_data <- data.frame(steps={},
                         date={},
                         interval={})
for(i in 1:length(X))
{
  new_data <- rbind(new_data, data.frame(steps=X[[i]]['steps'],
                                             date=X[[i]]['date'],
                                             interval=X[[i]]['interval']))
}
```

**Calculate a new mean, median, total, and create a histogram**
```{r}
tapply(new_data$steps, activity_data$weekday, mean, na.rm=T )
tapply(new_data$steps, activity_data$weekday, median, na.rm=T )
new_total <- aggregate(steps ~ date, new_data, sum)
hist(new_total$steps, xlab= "Total Steps Taken per Day", main="", col="blue", breaks = 10)
```

###Are there differences in activity patterns between weekdays and weekends?

```{r}
new_data['dayTypeInWeek'] = factor(sapply(new_data$date, function(x){ if (wday(x) == 6 | wday(x) == 7) { "weekend" } else { "weekday"} }))
avgStepdayTypeInWeek = aggregate(steps~interval + dayTypeInWeek, mean, data=new_data)

library(lattice)
xyplot( steps ~ interval | dayTypeInWeek, data = avgStepdayTypeInWeek, type="l", layout=c(1,2), xlab="Interval", ylab="Number of Steps")
```